### 2016-02-22 text/ms.Rmd
jessica edits - minor rewording and comments

\bigskip
\noindent
\textbf{Abstract:} Fishery managers must often reconcile conflicting estimates of population status and trend. One solution is to average predictions in an ensemble, but this approach ignores covariance between models and may not optimally reduce bias and improve accuracy. Superensemble models, commonly used in climate and weather forecasting, may provide a superior solution. Superensembles use predictions from multiple models as covariates in an additional model fit to known data. We evaluated the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of population status and trend. We fit four data-limited and widely applicable models of population and exploitation dynamics and combined the models' estimates of the mean and recent trend in biomass relative to the equilibrium biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an ensemble average and three superensembles: a linear model, a random forest, and a boosted regression tree. We built our models on a simulated dataset of 5760 stocks and tested them with cross-validation and against a database of `r ram_stocks_n` stock assessments. We found that ensemble methods substantially improved estimates of population status and trend. Random forest and boosted regression trees performed the best for estimating population status: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, rank-order correlation between predicted and true status improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements when predicting the trend of status and when applying the simulation-trained superensembles to catch data for global fish stocks. Ensemble methods can improve estimates of status and trends; however, they must be tested, formed from a diverse set of accurate models, and built on a dataset representative of the populations to which they are applied. *TODO: Still need to cut ~30 words*

@@

One solution is to take the average or weighted average of an ensemble of model predictions. Such ensembles are typically more accurate and less biased than individual model estimates and can integrate uncertainty in model structure, initial conditions, and parameter estimation [@dietterich2000; @araujo2007]. This approach forms the basis of many machine learning methods [e.g. @dietterich2000], has helped reconcile climate forecasts from dozens of models [e.g. @murphy2004; @tebaldi2007; @ipcc2013], and even improved early warning signs of malaria outbreaks [@thomson2006]. In ecology, ensemble methods are sometimes used to improve species distribution modelling [e.g. @araujo2007; @breiner2015] and indeed have been used to combine estimates of population status and trend [e.g. @brodziak2010].

Whereas averages or weighted averages may improve predictions over any single model, they may not optimally leverage available data. The best prediction does not necessarily lie in the middle of multiple model predictions, some models may perform better than others in certain conditions, and the covariance between models may contain information that can improve predictive accuracy. We can exploit these characteristics by using the predictions from a group of models as inputs into a separate statistical model. This technique, sometimes called superensemble modelling [@krishnamurti1999], is common in climate and weather forecasting [e.g. @yun2005; @mote2015]. The superensemble is fit to a training dataset where outcomes are well known and used to predict on a dataset of interest (Fig. \ref{didactic}). For example, @krishnamurti1999 combined predictions of wind and precipitation in Asian monsoons via a superensemble regression fit to observed data. Their superensemble was considerably more accurate than any individual prediction or weighted average of predictions.

In fisheries science, the commonly used operational models for determining status and trend of exploited fish populations are stock assessments, i.e., population models coupled to an observation model, that incorporate all appropriate data (e.g. catches, size and age distributions, surveys, and tagging information) to quantify values such as the biomass of a stock that can produce maximum sustainable yield ($B/B_\mathrm{MSY}$) [@hilborn1992]. However, the broad range of data required to conduct these stock assessments are not available for the majority of fish populations, including those of conservation concern and of economic interest to fisheries [@fao2014]. Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$ based on the limited data available for the majority of fish stocks: (1) a time series of the total biomass of catch and (2) a basic understanding of population productivity [e.g. @vasconcellos2005; @martell2013]. Recently, @rosenberg2014 investigated the performance of four data-limited models through a large-scale simulation experiment. Three of these models are based on Schaefer (logistic) biomass dynamics and one is an empirical model fitted to more data-rich stock assessment output. The four models frequently disagreed about population status, no one model had strong performance across all fish stocks, and some models performed better than others depending on circumstances (e.g. Fig. \ref{motivate}).

@@

We first developed and tested ensemble methods on a fully factorial simulated dataset [@rosenberg2014] with known status. These published simulated data included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of initial biomass depletion compared to carrying capacity: biomass at 100%, 70%, and 40% of carrying capacity; and four exploitation settings: (1) a constant exploitation rate, (2) an exploitation rate coupled with biomass to mimic an open-access single-species fishery, (3) a scenario where exploitation rate increased continuously, and (4) a 'roller-coaster' scenario where the exploitation rate increased and then decreased. Process noise (recruitment variability; i.e. unexplained variability in population dynamics) was introduced to the models at two magnitudes in log space, $N(0, 0.2^2)$ and $N(0, 0.6^2)$, and was either uncorrelated through time or had first-order autoregressive correlation of $0.6$. The simulation also included a scenario with $N(0, 0.2^2)$ measurement error around log(catch) and one scenario without measurement error. @rosenberg2014 ran ten iterations for each combination of factors adding stochastic draws of recruitment and catch-recording variability each time to generate a total of 5760 stocks. 
The simulation models were built in the package `FLR` [@kell2007, <http://flr-project.org>] for the statistical software R [@r2015].
Code to generate the simulations is available at <https://github.com/iagomosqueira/stockStatusFAO>.

@@

We fit four individual data-limited models to estimate $B/B_\mathrm{MSY}$. Three of the models are mechanistic and based generally on Schaefer biomass dynamics [@schaefer1954] of the form

@@

\noindent
where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus one year, $B_t$ represents biomass at time $t$, $r$ represents the intrinsic population growth rate, $B_0$ represents unfished biomass or carrying capacity $K$, and $C$ represents catch. The fourth model is an empirically derived model based on the RAM Legacy Stock Assessment Database. @rosenberg2014 provide a full background on these four methods 
and code to fit all the models is available in an accompanying package `datalimited` for the statistical software R TODO ADD URL. In summary:

@@

* *mPRM* (modified panel regression model) is a modified version of the panel-regression model from @costello2012. Unlike the other models, mPRM is empirical and not mechanistic --- it uses the RAM Legacy Stock Assessment database to fit a regression model to a series of characteristics of the catch time series and stock with stock-assessed $B/B_\mathrm{MSY}$ as the response. The model used in this paper is modified from the original --- it condenses the life-history categories into three categories to match the simulated dataset, removes the maximum catch predictor since the absolute catch in the simulated dataset is arbitrary, and does not implement the bias correction needed in @costello2012 since we do not derive estimates of median status across multiple stocks.

@@

<!--TODO Kristin and Mark: It might be a good idea to add a figure that illustrates the spectrum of ensemble/superensembles that are applied here and what is included (e.g., non-linear relationships, etc.). This will help the reader understand the progression in complexity of the ensemble approaches I think.-->
<!-- TODO deal with removed spectral -->
<!-- 
Superensemble models allow us to incorporate additional covariates and potentially leverage interactions between these covariates and individual model predictions. Additional covariates could be, for example, life-history characteristics, information on exploitation patterns, or statistical properties of the data. For simplicity, and since many possible covariates, such as fishing effort dynamics, would not be known for many real-world data-limited scenarios, we added only one set of additional covariates: spectral properties of the catch time series. Spectral analysis decomposes a time series into the frequency domain and provides a means of describing the cyclical shape of the catch series that is independent of time series length (except in affecting precision) and absolute magnitude of catch. We fit spectral models to the scaled catch time series (catch divided by maximum catch) with the `spec.ar` function in R and recorded representative short- and long-term spectral densities at frequencies of 0.20 and 0.05, which correspond to 5- and 20-year cycles.
-->

@@

We used the mean or slope of $B/B_\mathrm{MSY}$ as the response and the predictions from the individual models
<!--, and the spectral values, -->
as predictors in our superensemble models. When modelling mean $B/B_\mathrm{MSY}$ --- a ratio bounded at zero --- we fit the superensemble models in log space and exponentiated the predictions. For the estimates of $B/B_\mathrm{MSY}$ slope, which are not bounded at zero, we fit superensemble models on the natural untransformed scale.

@@

Predictive performance can be evaluated with metrics that represent a variety of modelling goals. For continuous response variables such as the mean and slope of population status, performance metrics often measure some form of bias, precision, accuracy (a combination of bias and precision), or the ability to correctly rank or correlate across populations [e.g. @walther2005]. Here, we measure proportional error, defined as $(\hat{b} - b)/\abs{b}$, where $\hat{b}$ and $b$ represent estimated and 'true' (or stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median proportional error to measure bias, median absolute proportional error to measure accuracy, and Spearman's rank-order correlation between predicted and 'true' values to measure the ability to correctly rank populations. When cross-validating against the RAM Legacy Stock Assessment database, we treated the estimates from these data-rich stock assessments as known without error. Thus, any error in the stock-assessment estimates of the mean or slope of $B/B_\mathrm{MSY}$ also contributes to our estimates of prediction error for each of the four data-limited models and the ensembles. We measured the ability for each model to correctly place stocks into categories of above or below $B/B_\mathrm{MSY} = 0.5$ at any decision threshold by plotting receiver operating characteristic curves; a threshold of $B/B_\mathrm{MSY} = 0.5$ is used to classify a stock as overfished in Australia and the United States [@daff2007; @hilborn2010]. Code to reproduce our analysis and an R package to fit the data-limited assessment models are available at <https://github.com/datalimited/...> (*Private while in review*).

@@

Certain conditions will make some ensemble models more effective than others. First, ensembles will be most effective when they are comprised of diverse individual models that choose different structural model forms, explore contrasting but plausible ranges of parameter values, or make uncorrelated errors [@ali1996; @dietterich2000; @tebaldi2007]. Such individual models would be expected to perform well in different conditions and an ensemble model can exploit the best predictive performance of each. Second, ensemble models will be most effective when they are not overfit to the training dataset. Cross-validation testing [@caruana2004; @hastie2009] and methods that are robust to overfitting such as random forests [@breiman2001], may help avoid overfitting ensemble models. We note that our simplest ensemble model, an average of individual model predictions, performed approximately as well as complex machine learning models when we trained our superensembles on the simulation dataset and tested them on a separate 'real' dataset (i.e. the RAM Legacy Stock Assessment database, Fig. \ref{performance}b). Third, ensemble models will be most effective when they are trained on data that are representative of the dataset of interest [@knutti2009; @weigel2010]. Cross-validation within a training dataset will provide an optimistically biased impression of predictive performance if the training dataset fundamentally differs from the dataset of interest [@hastie2009].

>>>>

\bigskip
\noindent
\textbf{Abstract:} Fishery managers must often reconcile conflicting estimates of population status and trend. One solution is to average predictions using an ensemble method, but this approach ignores covariance between models and may not optimally reduce bias and improve accuracy. Superensemble models, commonly used in climate and weather forecasting, may provide a superior solution. Superensembles use predictions from multiple models as covariates in an additional model. By training this superensemble model using known data, we can then use it to predict unknown situations. We evaluated the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of population status and trend for fisheries. We fit four widely applicable data-limited models that estimate stock biomass relative to the equilibrium biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$). We combined the models' estimates to build four ensemble methods: an ensemble average and three superensembles (a linear model, a random forest, and a boosted regression tree). We used the ensemble methods to predict both (i) fishery status and (ii) recent trends of $B/B_\mathrm{MSY}$. We trained our models on a simulated dataset of 5760 stocks and tested them with cross-validation and against a global database of `r ram_stocks_n` real stock assessments. We found that ensemble methods substantially improved estimates of population status and trend. Random forest and boosted regression trees performed the best for estimating population status: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, rank-order correlation between predicted and true status improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range` across models?, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements when predicting the trend of status and when applying the simulation-trained superensembles to catch data for global fish stocks. Ensemble methods can improve estimates of status and trends; however, they must be tested, formed from a diverse set of accurate models, and built on a dataset representative of the populations to which they are applied. *TODO: Still need to cut ~30 words*

@@

One solution is to take the average or weighted average of several model predictions, i.e. an ensemble. Such ensembles are typically more accurate and less biased than individual model estimates and can integrate uncertainty in model structure, initial conditions, and parameter estimation [@dietterich2000; @araujo2007]. <!--explain how this is possible.--> This approach forms the basis of many machine learning methods [e.g. @dietterich2000], has helped reconcile climate forecasts from dozens of models [e.g. @murphy2004; @tebaldi2007; @ipcc2013], and even improved early warning signs of malaria outbreaks [@thomson2006]. In ecology, ensemble methods are sometimes used to improve species distribution modelling [e.g. @araujo2007; @breiner2015] and indeed have been used to combine estimates of population status and trend [e.g. @brodziak2010].

Whereas averages or weighted averages of model estimates may improve predictions compared to a single model, they may not optimally leverage available data. The best prediction does not necessarily lie in the middle of multiple model predictions, some models may perform better than others in certain conditions, and the covariance between models may contain information that can improve predictive accuracy. <!--can you give an example here? e.g. a model may behave poorly at estimating low biomass. By training the superensbmeles models with real data, it can identify the pros and cons of each model and incorporate them into the estimates. .--> We can exploit these characteristics by using the predictions from a group of models as inputs into a separate statistical model. This technique, sometimes called superensemble modelling [@krishnamurti1999], is common in climate and weather forecasting [e.g. @yun2005; @mote2015]. The superensemble is fit to a training dataset where outcomes are well known and then used to predict on a dataset of interest (Fig. \ref{didactic}). For example, @krishnamurti1999 combined predictions of wind and precipitation in Asian monsoons via a superensemble regression fit to observed data. Their superensemble was considerably more accurate than any individual prediction or weighted average of predictions.

In fisheries science, the commonly used operational models for determining status and trend of exploited fish populations are stock assessments, i.e., population models coupled to an observation model, that incorporate all appropriate data (e.g. catches, size and age distributions, surveys, and tagging information) to quantify values such as the biomass of a stock that can produce maximum sustainable yield ($B_\mathrm{MSY}$) [@hilborn1992]. However, the broad range of data required to conduct these stock assessments are not available for the majority of fish populations, including those of conservation concern and of economic interest to fisheries [@fao2014]. Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$ based on the limited data available for the majority of fish stocks: (1) a time series of the total biomass of catch and (2) a basic understanding of population productivity [e.g. @vasconcellos2005; @martell2013]. Recently, @rosenberg2014 investigated the performance of four data-limited models through a large-scale simulation experiment. Three of these models are based on Schaefer (logistic) biomass dynamics and one is an empirical model fitted to more data-rich stock assessment output. The four models frequently disagreed about population status, no one model had strong performance across all fish stocks, and some models performed better than others depending on circumstances (e.g. Fig. \ref{motivate}).

@@

<!--
 Structure of methods: summary of methods
 1. indiviudla models
 2. ensemble models
 3. additional covariates
 4. data used to test and cross validation,
 5. testing model perforance. -->

We first developed and tested ensemble methods on a fully factorial simulated dataset of fisheries [@rosenberg2014] with known status. These simulated data included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of initial biomass depletion compared to carrying capacity: biomass at 100%, 70%, and 40% of carrying capacity; and four exploitation settings: (1) a constant exploitation rate, (2) an exploitation rate coupled with biomass to mimic an open-access single-species fishery, (3) a scenario where exploitation rate increased continuously, and (4) a 'roller-coaster' scenario where the exploitation rate increased and then decreased. Process noise (recruitment variability; i.e. unexplained variability in population dynamics) was introduced to the models at two magnitudes in log space, $N(0, 0.2^2)$ and $N(0, 0.6^2)$, and was either uncorrelated through time or had first-order autoregressive correlation of $0.6$. The simulation also included a scenario with $N(0, 0.2^2)$ measurement error around log(catch) and one scenario without measurement error. @rosenberg2014 ran ten iterations for each combination of factors adding stochastic draws of recruitment and catch-recording variability each time to generate a total of 5760 stocks. 
The simulation models were built in the `FLR` packages [@kell2007, <http://flr-project.org>] for the statistical software R [@r2015].
Code to generate the simulations is available at <https://github.com/iagomosqueira/stockStatusFAO>.

@@

We fit four individual data-limited models that use data on catch and basic life history parameters to estimate $B/B_\mathrm{MSY}$. <!-- explain why we choose these models not others? --> Three of the models are mechanistic and based generally on Schaefer biomass dynamics [@schaefer1954] of the form

@@

\noindent
where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus one year, $B_t$ represents biomass at time $t$, $r$ represents the intrinsic population growth rate, $B_0$ represents unfished biomass or carrying capacity $K$, and $C$ represents catch. The fourth model is an empirically derived model based on the RAM Legacy Stock Assessment Database. @rosenberg2014 provide a full background on these four methods 
and code to fit all the models is available in an accompanying package `datalimited` for the statistical software R TODO ADD URL. As described in the Introduction, the performance of these models at estimating $B/B_\mathrm{MSY}$ is variable. In summary:

@@

* *mPRM* (modified panel regression model) is a modified version of the panel-regression model from @costello2012. Unlike the other models, mPRM is empirical and not mechanistic --- it uses the RAM Legacy Stock Assessment database to fit a regression model to a series of characteristics of the catch time series and stock with stock-assessed $B/B_\mathrm{MSY}$ as the response. The model used in this paper is modified from the original --- it condenses the life-history categories into three categories to match the simulated dataset, removes the maximum catch predictor since the absolute catch in the simulated dataset is arbitrary, and does not implement the bias correction needed in @costello2012 since we do not derive estimates of median status across multiple stocks. <!--TODO is this last sentence necessary? -->

@@

<!--TODO Kristin and Mark: It might be a good idea to add a figure that illustrates the spectrum of ensemble/superensembles that are applied here and what is included (e.g., non-linear relationships, etc.). This will help the reader understand the progression in complexity of the ensemble approaches I think. 
Jess: probably add this below the section explaining superensembles-->
<!-- TODO deal with removed spectral -->
<!-- 
Superensemble models allow us to incorporate additional covariates and potentially leverage interactions between these covariates and individual model predictions. Additional covariates could be, for example, life-history characteristics, information on exploitation patterns, or statistical properties of the data. For simplicity, and since many possible covariates, such as fishing effort dynamics, would not be known for many real-world data-limited scenarios, we added only one set of additional covariates: spectral properties of the catch time series. Spectral analysis decomposes a time series into the frequency domain and provides a means of describing the cyclical shape of the catch series that is independent of time series length (except in affecting precision) and absolute magnitude of catch. We fit spectral models to the scaled catch time series (catch divided by maximum catch) with the `spec.ar` function in R and recorded representative short- and long-term spectral densities at frequencies of 0.20 and 0.05, which correspond to 5- and 20-year cycles.
-->

@@

We used the mean or slope of $B/B_\mathrm{MSY}$ as the response variable and the predictions from the individual models
<!--, and the spectral values, fix this explanation -->
as predictors in our superensemble models. When modelling mean $B/B_\mathrm{MSY}$ --- a ratio bounded at zero --- we fit the superensemble models in log space and exponentiated the predictions. For the estimates of $B/B_\mathrm{MSY}$ slope, which are not bounded at zero, we fit superensemble models on the natural untransformed scale.

@@

Predictive performance can be evaluated with metrics that represent a variety of modelling goals. For continuous response variables such as the mean and slope of population status, performance metrics often measure some form of bias, precision, accuracy (a combination of bias and precision), or the ability to correctly rank or correlate across populations [e.g. @walther2005]. Here, we measure proportional error, defined as $(\hat{b} - b)/\abs{b}$, where $\hat{b}$ and $b$ represent estimated and 'true' (or stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median proportional error to measure bias, median absolute proportional error to measure accuracy, and Spearman's rank-order correlation between predicted and 'true' values to measure the ability to correctly rank populations. When cross-validating against the RAM Legacy Stock Assessment database, we treated the estimates from these data-rich stock assessments as known without error. Thus, any error in the stock-assessment estimates of the mean or slope of $B/B_\mathrm{MSY}$ also contributes to our estimates of prediction error for each of the four data-limited models and the ensembles. We measured the ability for each model to correctly place stocks into categories of above or below $B/B_\mathrm{MSY} = 0.5$ at any decision threshold <!-- Jess: not sure what this means - at any decision threshold--> by plotting receiver operating characteristic curves; a threshold of $B/B_\mathrm{MSY} = 0.5$ is used to classify a stock as overfished in Australia and the United States [@daff2007; @hilborn2010]. Code to reproduce our analysis and an R package to fit the data-limited assessment models are available at <https://github.com/datalimited/...> (*Private while in review*).

@@

Certain conditions will make some ensemble models more effective than others. First, ensembles will be most effective when they are comprised of diverse individual models that choose different structural model forms, explore contrasting but plausible ranges of parameter values, or make uncorrelated errors [@ali1996; @dietterich2000; @tebaldi2007]. Such individual models would be expected to perform well in different conditions and an ensemble model can exploit the best predictive performance of each <!-- is this only true for superensembles?-->. Second, ensemble models will be most effective when they are not overfit to the training dataset. Cross-validation testing [@caruana2004; @hastie2009] and methods that are robust to overfitting such as random forests [@breiman2001], may help avoid overfitting ensemble models. We note that our simplest ensemble model, an average of individual model predictions, performed approximately as well as complex machine learning models when we trained our superensembles on the simulation dataset and tested them on a separate 'real' dataset (i.e. the RAM Legacy Stock Assessment database, Fig. \ref{performance}b). Third, ensemble models will be most effective when they are trained on data that are representative of the dataset of interest [@knutti2009; @weigel2010]. Cross-validation within a training dataset will provide an optimistically biased impression of predictive performance if the training dataset fundamentally differs from the dataset of interest [@hastie2009].
